---
title: "Regression Models Course Project"
author: "Kevin Huang"
date: "November 5, 2017"
output: pdf_document
---
# Executive Summary
In this project, we use model selection and linear regression to estimate the relationship between the transmission variable(*am*) and other independent variables, such as Weight(*wt*), Number of cylinders(*cyl*), Gross horsepower(*hp*), to figure out how the transmission will impact on *MPG*.  

We have concluded the following:  
1. *Manual* transmission has better *MPG* compare to *Automatic* transmission when we only use transmission along in the model. However, when we add in other variables, transmission has lower effect in terms of *MPG*.  
2. *MPG* will increase by 1.8 when the car is *manual* tranmission.

# Instrutions
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome).  
They are particularly interested in the following two questions:  
1. "Is an automatic or manual transmission better for MPG"  
2. "Quantify the MPG difference between automatic and manual transmissions"

# Data Description
The data set *mtcars* contains a data frame with 32 observations on 11 variables.

[, 1]	 mpg	 Miles/(US) gallon  
[, 2]	 cyl	 Number of cylinders  
[, 3]	 disp	 Displacement (cu.in.)  
[, 4]	 hp	 Gross horsepower  
[, 5]	 drat	 Rear axle ratio  
[, 6]	 wt	 Weight (1000 lbs)  
[, 7]	 qsec	 1/4 mile time  
[, 8]	 vs	 V/S  
[, 9]	 am	 Transmission (0 = automatic, 1 = manual)  
[,10]	 gear	 Number of forward gears  
[,11]	 carb	 Number of carburetors  

# Data Processing, Transformation, and Exploratory data analysis
We load the data into R, and convert some of the variables to factors.

```{r}
data(mtcars)
df <- mtcars
names(df)
dim(df)
head(df)
str(df)
df$cyl <- as.factor(df$cyl)
df$vs <- as.factor(df$vs)
df$am <- as.factor(df$am)
df$gear <- as.factor(df$gear)
df$carb <- as.factor(df$carb)
summary(df)
```

# Inference
Before we do the model selection, we perform a t-test to test if there is significant difference in mean *mpg* between *automatic* and *manaual* transmission.

```{r}
result <- t.test(mpg ~ am, df)
```

The result of the t-test shows the p-value is `r round(result$p.v, 3)`; therefore, it is significantly different in the mean of *automatic* and *manual.* The mean *mpg* for *automatic* transmission is `r round(result$est[1], 3)`, and the mean *mpg* for *manual* transmission is `r round(result$est[2], 3)`.

We can also see the boxplot in appendix 1 comparing the means *mpg* for *manual* and *automatic* transmission.

# Regression Analysis and Model Selection
### Simple Linear Regression

First off, we fit a initial model which *MPG* as outcome and *am* as the only predictor
```{r}
iniModel <- lm(mpg ~ am, df)
summary(iniModel)
```

The p-value is almost 0, whcih tells us the variable *am* is significant. However, the R-squares is `r round(summary(iniModel)$r.sq, 2)`, which means only `r round(summary(iniModel)$r.sq, 2) * 100`% of the variance is explianed by this model. Therefore, we will have to fit other models that includes significant variables to explian the variance.

### Model Selection
By looking at pairs plot (appendix 2) and the correlations between *MPG* and variables (appendix 3), I choose *cyl*, *disp*, *hp*, *wt* along with *am* to fit more models because they are highly correlated to *MPG.* I will be using nested model testing to find the significant variables.

```{r}
fit1 <- lm(mpg ~ am + wt, df)
fit2 <- lm(mpg ~ am + wt + cyl, df)
fit3 <- lm(mpg ~ am + wt + cyl + disp, df)
fit4 <- lm(mpg ~ am + wt + cyl + disp + hp, df)
anova.test.1 <- anova(iniModel,fit1, fit2, fit3, fit4)

fit4_rm_disp <- lm(mpg ~ am + wt + cyl + hp, df)
anova.test.2 <- anova(iniModel,fit1, fit2, fit4_rm_disp)

bestModel <- lm(mpg ~ am + wt + cyl + hp, df)
```
I put *am* as first predictors and followed by the order of correlation to *MPG*. Then I run the anova analysis (appendix 4). From model 3 and model 4, we can see that by adding *disp* does not have much impact to the model, but model 5 is significant, meaning adding *hp* is significant. Therefore, I remove disp and run anova again (appendix 5).

The anova analysis suggests that the 4th model, removed disp, is significant. We have the best model, *mpg ~ am + wt + cyl + hp*. 

### Multiple Linear Regression
```{r}
best <- summary(bestModel)
best
```
The p-values are small, the R-square is `r round(best$r.s, 3)`, the coefficient for *am* is `r round(best$coe[2], 2)` which suggests that driving *manual* transmission car will increase *mpg* by `r round(best$coe[2], 2)` than *automatic*.

# Residual and Diagnostics
We will use the residual plot to diagnosis if there is any outlier that affect the model.
```{r}
par(mfrow = c(2,2))
plot(bestModel)
```

From the above plot, we have a few observations,  
1. The *Residuals vs Fitted* plot does not appear a linear relationship.  
2. The *QQ plot* appears a little bit of tail, but it is normal overall. Therefore, it suggests that the residuals are normally distributed. 
3. The *Scale Location* plot appears a horizontal line, again, it suggests the residuals are spread equally along the ranges of predictors, meaning homoscedasticity.  
4. The *Residual vs Leverage* plot appears there is no influtial case in the model.

# Conclusion
1. After the model selection, our best model is *lm(mpg ~ am + wt + cyl + hp, df)*.
2. After performing the residuals analysis, we can be sure our model is correct since the residuals are normally distributed, and there is no influtial case in the model.
3. *MPG* will increase 1.8 when choosing *manual* transmission over *automatic*.

# Appendix
### Appendix 1
```{r, fig.height=3}
library(ggplot2)
ggplot(df, aes(x = am, y = mpg, group = am)) + 
    geom_boxplot(aes(fill = am)) + 
    ggtitle("Automatic v.s Manual Transmission") +
    xlab("Transmission Type, 0 = Automatic, 1 = Manual") + ylab("MPG") +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_discrete(name="Transmission Type", 
                        breaks=c("0", "1"), 
                        labels=c("Automatic", "Manual"))
```

### Appendix 2
```{r}
pairs(df)
```


### Appendix 3
```{r}
suppressPackageStartupMessages(library(Hmisc))
cor <- rcorr(as.matrix(df))
cor$r[1,]
```

### Appendix 4
```{r}
anova.test.1
```
### Appendix 5
```{r}
anova.test.2
```
